### 2016-05-16 admin/coopTimeline/timeline.tex
init!

    \textit{We don't need to be ``Internet--scale'' right off the bat; the big challenge will be figuring out how to make interactions work in a digital medium; we'll tackle the ``interactions at scale'' issue later.}
    \item Hashing out norms ---
    bylaws, policies, etc\dots
    \item Figuring out system specifications ---
    Defining how an automatic system should behave \& what interactions it should make available
    (e.g. how much information about workers customers should be able to see)

@@

    \textit{The goal of this 3--month phase is to establish a running track of the cooperative working (both in the sense of getting work done and functioning properly).}
  \end{itemize}
\section*{May: Write}
  \begin{itemize}
    \item I'll want to write about the findings of this
    (hopefully) successfully--running cooperative work group.
    With the deadline for \href{http://cscw.acm.org/2016/index.php}{Computer--Supported Cooperative--Work}
    (A Human--Computer Interaction conference) in late May,
    I'll want to have ample time to write up my notes.
  \end{itemize}
\section*{Later: Scale}

>>>>

    \textit{We don't need to be ``Internet--scale'' right off the bat; the big challenge will be figuring out how to make interactions work in a digital medium; we'll tackle the ``interactions at scale'' issue later.}
    \item Hashing out norms ---
    bylaws, policies, etc\dots
    \item Figuring out system specifications ---
    Defining how an automatic system should behave \& what interactions it should make available
    (e.g. how much information customers should be able to see)

@@

    \textit{The goal of this 3--month phase is to establish a running track of the cooperative working (both in the sense of getting work done and functioning properly).}
  \end{itemize}
\section*{May: Write}
  \begin{itemize}
    \item I'll want to write about the findings of this
    (hopefully) successfully--running cooperative work group.
    With the deadline for \href{http://cscw.acm.org/2016/index.php}{Computer--Supported Cooperative--Work}
    (A Human--Computer Interaction conference) in late May,
    I'll want to have ample time to write up my notes.
  \end{itemize}
\section*{Later: Scale}
  \begin{itemize}
    \item First to around 50 workers, then to around 100, then to around 150 or 200.
    \item \textit{Each milestone represents a different \textbf{kind} of group --- a group of 10 operates differently from a group of 50; a 50--person group operates differently from a 100--person group; etc\dots}
  \end{itemize}

### 2016-05-16 proposals/nsfGRFP/research_statement.tex
init!

\begin{document}
  \begin{center}
  \large{\bf Crowd--Operated Gig Markets and Crowdsourcing}
  % Ali Alkhatib
  \end{center}

\sectitle{introduction}\\
In the past several years, researchers have watched new forms of labor,
ranging from information work like Amazon Mechanical Turk (AMT)
to offline gigs like Uber and TaskRabbit emerge and balloon into industries now worth billions.
While workers in these markets point out that the transient ``gig work'' these markets enable
allow them to participate in the market when they might have been unable to do so before,
for instance due to medical conditions preventing them from making long--term commitments,
or competing obligations such as parenthood or caregiving.

On the other hand, this ``gig economy'' has been criticized for commoditizing
and ultimately marginalizing workers
\cite{uberAlgorithm}.
Efforts have attempted to make these markets more rewarding, or at least more bearable,
for workers
\cite{turkopticon,dynamo},
but we still see no fundamentally differentiated labor markets for workers;
virtually all of the existing markets trade one manager for another, maintaining an adversarial role between laborers and their managers.

Workers have faced issues like these for more than a century,
but they bear relevance to the HCI community broadly for two major reasons:
\begin{inlinelist}
  \item the features of digital systems allow researchers to have an outsize impacts on workers,
  whereas historically markets have been fractured \& untenable,
  and
  % --- or emulate their characteristics ---
  \item as researchers of technical systems and social interactions,
  researchers have a prerogative to attempt to steer systems toward more beneficial outcomes.
  \textit{researchers, broad impact, etc\dots}
\end{inlinelist}

\sectitle{hypothesis}\\
I propose to investigate the viability of a technologically--mediated,
worker--run market as an alternative to existing systems.
I hypothesize that
such a market can compete effectively,
and that workers and consumers would benefit more, % than under the status quo,
for a number of reasons.
% as a result of community--determined norms and rules.
% and that it would benefit workers \& consumers due to the equitable relationships that it inherently fosters.
For workers,
\begin{inlinelist} % \itemsep 0pt \parskip 0pt
  \item communally determined rules and norms are perceived as, and perhaps are, fairer, and
  \item a protocol for labor markets would allow markets to communicate things like worker reputation.
\end{inlinelist}
Customers, for their parts, would be able
% \begin{inlinelist} % \itemsep 0pt \parskip 0pt
  % \item to discriminate between markets on various criteria, and
  to leverage market competition, minimizing predatory pricing models.
% \end{inlinelist}

@@

The challenge of a worker--run market seems to be sociocultural.
The emergence, communication, and enforcement of norms
add a dimension of complexity to the individual cases of collective action represented by the study of individual events.
Engineering instances of collective action is a qualitatively different challenge from
fostering a community of ongoing collective action
\cite{russell1982collective}.
% (e.g. \cite{catalyst,dynamo}).

@@

The most immediate goal is to implement a system consisting of
the technical requirements of gig labor markets described before
(factors such as scheduling, dispatch, \& payments).
This system would resemble a mobile application,
similar in appearance to existing labor platforms,
but importantly differentiated by affordances for community governance:
\begin{inlinelist}
  \item forums for discourse,
  \item voting systems for referendums, and
  \item bylaws written and approved by the community approval.
\end{inlinelist}

The Fair Care Labs plans to deploy this system in two locales
--- the San Francisco Bay Area and the Seattle area ---
in order to allow me to provide high--level guidance
and to make course--corrections as necessary.
This should ameliorate concerns both about funding the system
and sourcing workers;
my involvement is the extent to which I will be able to contribute as a researcher.

@@

% In this area, I am relatively uniquely qualified.
The ethnographic and design work I've done at Microsoft Research (MSR),
my work at Stanford University (Dynamo
\cite{dynamo}),
and my background and training as an Anthropologist
make me ideally skilled to pursue this area.
I have demonstrated that I can build systems,
and crucially I have proven experience in the qualitative,
nebulous fieldwork necessary for this research.

@@

\renewcommand\refname{\sectitle{references}}
% % REFERENCES FORMAT
% % References must be the same font size as other body text.
% \bibliographystyle{SIGCHI-Reference-Format}
\bibliographystyle{acm}
{\footnotesize
\bibliography{references}
}

>>>>

\begin{document}
  \begin{center}
  \large{\bf Worker--Managed Crowdsource Labor Markets}
  % Ali Alkhatib
  \end{center}

\sectitle{introduction}\\
In the past several years, researchers have watched new forms of labor
--- exemplified by labor markets like Amazon Mechanical Turk (AMT), Uber, and TaskRabbit ---
emerge and balloon into billion--dollar industries.
%
While these markets allow people to engage in the workforce in novel ways,
this ``gig economy'' has been criticized for commoditizing
and marginalizing workers
\cite{uberAlgorithm}.

On Turkopticon \cite{turkopticon} and in my prior work on Dynamo \cite{dynamo},
we attempted to make these markets more rewarding for workers,
but no labor markets to date place workers first;
existing markets trade one manager for another, maintaining an adversarial role between laborers and their managers.

Workers have faced issues like these for more than a century,
but they bear relevance to the HCI community broadly for two major reasons:
\begin{inlinelist}
  \item the features of digital systems allow researchers to have an outsize impacts on workers,
  whereas historically markets have been fractured \& untenable,
  and
  % --- or emulate their characteristics ---
  \item as researchers of technical systems and social interactions,
  researchers have a prerogative to attempt to steer systems toward more beneficial outcomes,
  especially given the growth of this form of labor and the potential implications for the future of work.
\end{inlinelist}

\sectitle{hypothesis}\\
I propose to investigate the viability of a technologically--mediated,
worker--run market as an alternative to existing systems.
I hypothesize that
such a market can compete effectively,
and that workers and consumers would benefit more, % than under the status quo,
for a number of reasons.
% as a result of community--determined norms and rules.
% and that it would benefit workers \& consumers due to the equitable relationships that it inherently fosters.
For workers,
\begin{inlinelist} % \itemsep 0pt \parskip 0pt
  \item communally determined rules and norms are perceived as, and perhaps are, fairer, and
  \item a protocol for labor markets would allow markets to communicate things like worker reputation.
\end{inlinelist}
Customers, for their parts, would be able
to leverage market competition, minimizing predatory pricing models.

@@

No; we're finding that the challenges to worker--centric markets are social.
Research has shown that crowds can be directed
and can make tactical decisions on an ad hoc basis,
but collective governance and policy--making,
especially in competitive economic settings, remains an open question.
% The emergence, communication, and enforcement of norms
The coalescence of culture
adds a dimension of complexity to the individual cases of collective action represented by the study of individual events.
Engineering instances of collective action is a qualitatively different challenge from
fostering a community of ongoing collective action
\cite{russell1982collective}.
% (e.g. \cite{catalyst,dynamo}).

@@

My immediate task is to create a labor market which illustrates the characteristics described
--- technical features like scheduling, dispatch, \& payments, as well as
the necessary affordances for collective governance:
% The most immediate goal is to implement a system consisting of
% the technical requirements of gig labor markets described before
% (factors such as scheduling, dispatch, \& payments).
% I will develop a mobile application,
% similar in appearance to existing labor platforms,
% but importantly differentiated by affordances for community governance:
% \begin{inlinelist}
  % \item
  forums for discourse,
  % \item
  voting systems for referendums, and
  % \item
  communally written bylaws. %  written and approved by the community.
% \end{inlinelist}
These tools will be the starting point for a worker--run crowd labor market,
upon which more affordances will be designed and implemented as needed.

The Fair Care Labs plans to continue the relationship that began over the summer while I was affiliated with Microsoft Research,
deploying this system in the San Francisco Bay Area and the Seattle area
in order to allow me to provide high--level guidance
and to make any necessary course--corrections.
This should ameliorate concerns both about funding the system
and sourcing workers;
my involvement is the extent to which I will be able to contribute as a researcher.

@@

% In this area, I am relatively uniquely qualified.
The ethnographic and design work I've done with various labor advocacy groups,
my prior work on Dynamo
\cite{dynamo},
and my background and training as an Anthropologist
make me ideally skilled to pursue this line of research.
I have demonstrated that I can build systems,
and crucially I have proven experience in the qualitative,
sometimes nebulous fieldwork necessary for this research.

@@

\renewcommand\refname{\sectitle{references}}
% % REFERENCES FORMAT
% % References must be the same font size as other body text.
% \bibliographystyle{SIGCHI-Reference-Format}
\bibliographystyle{acm}
{\footnotesize
\bibliography{../references.bib}
}
